
# METHOD

To compare the rhetorical strategies of Ronald Reagan and Donald Trump Readability was evaluated through two established indices: the Dale–Chall (DC) and the Flesch–Kincaid (FK) scores, each emphasizing different aspects of textual complexity. Together, these metrics allow us to examine not only how varied the presidents’ vocabularies were but also how accessible or demanding their speeches appeared to general audiences.

## Import the libraries

To Perfom the Natural Language processing we need to  import the necessary libraries
```{r setup, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
#install.packages(c("xml2", "tidyverse", "quanteda"))
#tinytex::install_tinytex()
library(xml2)
library(rmarkdown)
library(textstem)
library(spacyr)
library(stringr)
library(readr)
library(quanteda)
library(readtext)
library(quanteda.textstats)
library(tokenizers)
library(tidytext)
library(tibble)
library(dplyr)
library(ggplot2)

```

We set the working directory to the folder containing our XML files: cleaned_reagan_speeches.xml and cleaned_trump_speeches.xml
```{r}
setwd("C:/Users/leo_m/Desktop/Corpus_Analysis/Final_Project/Reagan_Trump")
```

## XML to Data Frame

We transformed the XML files into a data frame, converting the XML nodes into table columns using the xml2 library.
```{r}
parse_speeches <- function(file_path, speaker) {
  # Read XML file
  xml_data <- read_xml(file_path)
  
  # Extract all speech nodes
  speeches <- xml_find_all(xml_data, "//speech")
  
  # Extract each node as a vector
  titles <- xml_text(xml_find_all(speeches, "title"))
  dates <- xml_text(xml_find_all(speeches, "date"))
  links <- xml_text(xml_find_all(speeches, "link"))
  texts <- xml_text(xml_find_all(speeches, "text"))
  
  # Create a data frame
  df <- tibble(
    speaker = speaker,
    title = titles,
    date = dates,
    link = links,
    text = texts
  )
  
  return(df)
}
```

### Create a Data Frame of the speeches
We create a data frame for each president and the we combine both in one data frame speeches_df
```{r}
# Parse Reagan speeches
reagan_df <- parse_speeches("cleaned_reagan_speeches.xml", "Reagan")

# Parse Trump speeches
trump_df <- parse_speeches("cleaned_trump_speeches.xml", "Trump")

# Combine into one data frame
speeches_df <- bind_rows(reagan_df, trump_df)

#Preview of the data frame
head(speeches_df)
dim(speeches_df)

# Count the number of speeches for each speaker
speech_counts <- speeches_df %>%
  group_by(speaker) %>%
  summarise(count = n())
print(speech_counts)
```

### Balancing same number of speeches for each President

After scraping, there were more Donald Trump speeches available in the American Presidency Project. To ensure a balanced dataset for comparison, we randomly selected 139 speeches for Donald Trump using sample_n(), matching the number of speeches available for Ronald Reagan. We also check for any duplicates 
```{r}
# Set a seed for reproducibility
set.seed(123)  # Ensures consistent random sampling

# Filter Reagan speeches (keep all) and sample 139 Trump speeches
balanced_speeches_df <- speeches_df %>%
  filter(speaker == "Reagan") %>%
  bind_rows(
    speeches_df %>%
      filter(speaker == "Trump") %>%
      sample_n(size = 139)
  )
# Check the number of speeches per speaker
balanced_counts <- balanced_speeches_df %>%
  group_by(speaker) %>%
  summarise(count = n())
print(balanced_counts)

# Inspect the data frame (Show 5 random rows)

print(balanced_speeches_df[sample(nrow(balanced_speeches_df), 5), ])


# Check for duplicate titles
duplicate_titles <- balanced_speeches_df %>%
  group_by(title) %>%
  summarise(count = n()) %>%
  filter(count > 1)
print(duplicate_titles)
```

Even when the titles appear similar, the dates differ, indicating that the events were the same but the speeches were delivered in different contexts.


### Saving XML file for selected 139 speeches of Donald Trump
```{r}
# Create root node
root <- xml_new_root("speeches")

# Loop over speeches
for (i in seq_len(nrow(balanced_speeches_df))) {

  speech_node <- xml_add_child(root, "speech")

  xml_add_child(speech_node, "speaker",
                balanced_speeches_df$speaker[i])

  xml_add_child(speech_node, "title",
                balanced_speeches_df$title[i])

  xml_add_child(speech_node, "date",
                balanced_speeches_df$date[i])

  xml_add_child(speech_node, "text",
                balanced_speeches_df$text[i])
  
  

}
write_xml(root, "trump_139_selected_speeches.xml")  

```



Even when the titles appear similar, the dates differ, indicating that the events were the same but the speeches were delivered in different contexts.



### Creating a Quanteda Corpus

We first added a unique ID to each speech using mutate() and sprintf("%03d", row_number()), creating IDs like 001, 002, etc. Then, we built a quanteda corpus with corpus(), specifying the text column (text_field = "text") and the document ID (docid_field = "unique_id"). This allows efficient text analysis. Finally, we checked the corpus structure with print(), summary(), and head(docnames()) to confirm the documents were correctly labeled.

```{r}

balanced_speeches_df <- balanced_speeches_df %>%
  mutate(unique_id = sprintf("%03d", row_number()))
# Create a quanteda corpus
speeches_corpus <- corpus(balanced_speeches_df, 
                          text_field = "text",
                          docid_field = "unique_id")

```

```{r}
# Step 3: Check the corpus structure
print(speeches_corpus)
speeches_corpus_df <- summary(speeches_corpus)
head(docnames(speeches_corpus))
```


Then We generated a summary of the corpus using summary(speeches_corpus, n = 5), which shows the first five documents. Then, we tokenized the corpus with quanteda::tokens(), splitting the text into individual words for analysis.

We extracted the document names with docnames() and the speakers from the document variables using docvars(), storing them in speakers_vec and naming them according to their document IDs. This links each speech to its speaker.



```{r}
# Summary of the corpus
summary(speeches_corpus, n = 5)  # Shows first 5 documents

speeches_tokens <- quanteda:: tokens(speeches_corpus)

docnames_vec <- docnames(speeches_tokens)
speakers_vec <- docvars(speeches_tokens, "speaker")
names(speakers_vec) <- docnames_vec



```



### Overall Token/sentences counts


```{r}



# Tokenize the corpus (word tokens)
speeches_tokens <- quanteda::tokens(
  speeches_corpus,
  remove_punct = TRUE
)


total_tokens <- sum(ntoken(speeches_tokens))
print(total_tokens)

# Sentence tokenization 
speeches_sentences <- quanteda::tokens(
  speeches_corpus,
  what = "sentence"  ### tokenize by sentences units!!!!
)

total_sentences <- sum(ntoken(speeches_sentences))
print(total_sentences)

# Tokens and Sentences per document
token_counts <- ntoken(speeches_tokens)
sentence_counts <- ntoken(speeches_sentences)

# Document-level metadata
doc_info <- docvars(speeches_corpus)

# Combine counts with metadata
count_data <- data.frame(
  speaker   = doc_info$speaker,
  tokens    = token_counts,
  sentences = sentence_counts
)

# Aggregate totals by speaker
aggregate_stats <- count_data %>%
  group_by(speaker) %>%
  summarise(
    total_tokens     = sum(tokens),
    avg_tokens       = mean(tokens),
    sd_tokens        = sd(tokens),
    total_sentences  = sum(sentences),
    avg_sentences    = mean(sentences),
    sd_sentences     = sd(sentences),
    n_speeches       = n()
  )

print(aggregate_stats)



```










### Creating a Balanced Chunked Corpus of 30 sentences

When readability is calculated using token-based chunks instead of full sentences, the original sentence structures are disrupted. This can distort measures of , average sentence length, and other readability indicators, because the coherent units of meaning—sentences—are broken into arbitrary segments. By using equal samples of full sentences (e.g., 30 per sample), we preserve the natural syntactic and rhetorical structures, ensuring that readability scores accurately reflect the complexity and organization of the text.

```{r}


set.seed(123)

# =====================================================
# 1️⃣ Tokenizar todos los discursos por oraciones
# =====================================================
sentences_tokens <- tokens(speeches_corpus, what = "sentence")
all_sentences <- unlist(sentences_tokens)

# Crear data frame de oraciones con metadata
sentences_df <- data.frame(
  sentence_id  = seq_along(all_sentences),
  text         = all_sentences,
  original_doc = rep(docnames(speeches_corpus), ntoken(sentences_tokens)),
  stringsAsFactors = FALSE
)

# Añadir speaker
sentences_df <- sentences_df %>%
  left_join(
    data.frame(original_doc = docnames(speeches_corpus), 
               speaker = docvars(speeches_corpus, "speaker")),
    by = "original_doc"
  )

# =====================================================
# 2️⃣ Crear chunks de 30 oraciones
# =====================================================
CHUNK_SIZE <- 30

sentences_df <- sentences_df %>%
  group_by(original_doc) %>%
  mutate(chunk_id = paste0(original_doc, "-", ceiling(row_number() / CHUNK_SIZE))) %>%
  ungroup()

# Combinar las oraciones de cada chunk
chunked_df <- sentences_df %>%
  group_by(chunk_id, speaker) %>%
  summarise(
    text       = paste(text, collapse = " "),
    nsentences = n(),
    .groups = "drop"
  )

# =====================================================
# 3️⃣ Eliminar chunks con menos de 30 oraciones
# =====================================================
chunked_df_clean <- chunked_df %>%
  filter(nsentences >= CHUNK_SIZE)

# =====================================================
# 4️⃣ Calcular tokens por chunk
# =====================================================
chunked_toks <- tokens(chunked_df_clean$text,
                       remove_punct   = TRUE,
                       remove_numbers = TRUE,
                       remove_symbols = TRUE)

chunked_df_clean$ntokens <- ntoken(chunked_toks)

# =====================================================
# 5️⃣ Balancear speakers (Trump → Reagan)
# =====================================================
# Número de chunks de Reagan
n_reagan <- chunked_df_clean %>% filter(speaker == "Reagan") %>% nrow()

# Separar por speaker
chunks_list <- chunked_df_clean %>% group_split(speaker)
names(chunks_list) <- sapply(chunks_list, function(df) unique(df$speaker))

# Tomar todos los de Reagan y samplear Trump
balanced_chunks <- bind_rows(
  chunks_list[["Reagan"]],
  slice_sample(chunks_list[["Trump"]], n = n_reagan, replace = FALSE)
)

# =====================================================
# 6️⃣ Crear data frame final del corpus
# =====================================================
corpus_df <- balanced_chunks %>%
  select(
    chunk_id,
    speaker,
    text,
    ntokens,
    nsentences
  )

# =====================================================
# 7️⃣ Verificación rápida
# =====================================================
# Número de chunks por speaker
corpus_df %>% count(speaker, name = "n_chunks")

# Resumen de tokens y oraciones
summary(corpus_df$ntokens)
summary(corpus_df$nsentences)


```
### Comparison Sentence vs Token lenghts
```{r}


summary_by_speaker <- corpus_df %>%
  group_by(speaker) %>%
  summarise(
    n_chunks       = n(),
    total_tokens   = sum(ntokens),
    total_sentences= sum(nsentences),
    .groups = "drop"
  )

print(summary_by_speaker)


```

### Distribution Sentences-Tokens by Speaker
```{r}


summary_by_speaker <- corpus_df %>%
  group_by(speaker) %>%
  summarise(
    n_chunks       = n(),
    total_tokens   = sum(ntokens),
    total_sentences= sum(nsentences),
    .groups = "drop"
  )

print(summary_by_speaker)


```

## Calculating Readability Metrics: 

This code computes several readability metrics—including Flesch, Flesch-Kincaid, Dale-Chall, SMOG, and Gunning Fog—for each text chunk in the corpus. By aggregating these metrics at the speaker level, we can compare the average readability, variation, and extremes of speech across different speakers. Maintaining sentence-based chunks ensures that readability scores reflect true syntactic and lexical complexity rather than artifacts of chunking.
```{r}




# ===============================
# 1 Compute readability metrics (using FOG for Gunning Fog)
# ===============================
readability_metrics <- textstat_readability(
  corpus_df$text,
  measure = c("Flesch", "Flesch.Kincaid", "Dale.Chall", "SMOG", "FOG")
)

# Combine with corpus_df
corpus_df <- bind_cols(corpus_df, readability_metrics)

# Optional: rename columns for simplicity
corpus_df <- corpus_df %>%
  rename(
    Flesch        = Flesch,
    FleschKincaid = `Flesch.Kincaid`,
    DaleChall     = `Dale.Chall`,
    GunningFog    = FOG
  )

# ===============================
# 2 Quick check
# ===============================
head(corpus_df %>% 
       select(chunk_id, speaker, ntokens, nsentences, 
              Flesch, FleschKincaid, DaleChall, SMOG, GunningFog))

# ===============================
# 3 Summary statistics by speaker (including Gunning Fog)
# ===============================
readability_stats <- corpus_df %>%
  group_by(speaker) %>%
  summarise(
    mean_Flesch      = mean(Flesch, na.rm = TRUE),
    sd_Flesch        = sd(Flesch, na.rm = TRUE),
    min_Flesch       = min(Flesch, na.rm = TRUE),
    max_Flesch       = max(Flesch, na.rm = TRUE),
    
    mean_FK          = mean(FleschKincaid, na.rm = TRUE),
    sd_FK            = sd(FleschKincaid, na.rm = TRUE),
    min_FK           = min(FleschKincaid, na.rm = TRUE),
    max_FK           = max(FleschKincaid, na.rm = TRUE),
    
    mean_DC          = mean(DaleChall, na.rm = TRUE),
    sd_DC            = sd(DaleChall, na.rm = TRUE),
    min_DC           = min(DaleChall, na.rm = TRUE),
    max_DC           = max(DaleChall, na.rm = TRUE),
    
    mean_SMOG        = mean(SMOG, na.rm = TRUE),
    sd_SMOG          = sd(SMOG, na.rm = TRUE),
    min_SMOG         = min(SMOG, na.rm = TRUE),
    max_SMOG         = max(SMOG, na.rm = TRUE),
    
    mean_GunningFog  = mean(GunningFog, na.rm = TRUE),
    sd_GunningFog    = sd(GunningFog, na.rm = TRUE),
    min_GunningFog   = min(GunningFog, na.rm = TRUE),
    max_GunningFog   = max(GunningFog, na.rm = TRUE),
    
    n_chunks         = n()
  )

readability_stats


```

### Calculating t.test() Readability Scores
```{r}

metrics <- c("Flesch", "FleschKincaid", "DaleChall", "SMOG", "GunningFog")

# Create a list to store results
t_test_results <- lapply(metrics, function(metric) {
  t.test(corpus_df[[metric]] ~ corpus_df$speaker)
})

# Name the list by metric
names(t_test_results) <- metrics

# Print all results
t_test_results

# Or view in a cleaner format
for(metric in metrics){
  cat("\n---", metric, "---\n")
  print(t_test_results[[metric]])
}


```

## Calculating Correlation Between Readability Scores
```{r}
library(dplyr)
library(corrplot)

# ===============================
# 1️⃣ Select readability metrics
# ===============================
correlation_df <- corpus_df %>%
  select(
    Flesch,
    FleschKincaid,
    SMOG,
    GunningFog,
    DaleChall
  )

# ===============================
# 2️⃣ Correlation matrix (Pearson)
# ===============================
cor_matrix <- cor(
  correlation_df,
  use = "pairwise.complete.obs",
  method = "pearson"
)

# Print correlation matrix
print(cor_matrix)

# ===============================
# 3️⃣ Visualization
# ===============================
corrplot(
  cor_matrix,
  method = "color",
  type = "upper",
  addCoef.col = "black",
  tl.cex = 0.9,
  number.cex = 0.8,
  title = "Correlation Matrix: Readability Metrics",
  mar = c(0, 0, 1, 0)
)


```
### Visualization Readability Scores 
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

# -----------------------------
# 1. Prepare data for plotting
# -----------------------------
plot_df <- corpus_df %>%
  select(speaker, Flesch, FleschKincaid) %>%  # just select existing columns
  pivot_longer(
    cols = c(Flesch, FleschKincaid),
    names_to = "metric",
    values_to = "value"
  )

# -----------------------------
# 2. Plot boxplots
# -----------------------------
ggplot(plot_df, aes(x = speaker, y = value, fill = speaker)) +
  geom_boxplot(alpha = 0.7, outlier.color = "black", outlier.shape = 16) +
  facet_wrap(~ metric, ncol = 2, scales = "free_y") +
  labs(
    title = "Distribution of Readability Scores by Speaker",
    x = "Speaker",
    y = "Score"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 30, hjust = 1)
  )
```

```{r}
ggplot(corpus_df, aes(x = speaker, y = SMOG, fill = speaker)) +
  geom_boxplot(alpha = 0.7) +
  labs(
    title = "Distribution of SMOG Scores by Speaker",
    x = "Speaker",
    y = "SMOG Score"
  ) +
  theme_minimal() +
  theme(
    text = element_text(size = 12),
    legend.position = "none"
  )
```
```{r}
ggplot(corpus_df, aes(x = speaker, y = GunningFog, fill = speaker)) +
  geom_boxplot(alpha = 0.7) +
  labs(
    title = "Distribution of Gunning Fog Indexes by Speaker",
    x = "Speaker",
    y = "Gunning Fog"
  ) +
  theme_minimal() +
  theme(
    text = element_text(size = 12),
    legend.position = "none"
  )

```

```{r}
library(ggplot2)

ggplot(corpus_df, aes(x = speaker, y = DaleChall, fill = speaker)) +
  geom_boxplot(alpha = 0.7) +
  labs(
    title = "Distribution of Dale–Chall Readability Scores by Speaker",
    x = "Speaker",
    y = "Dale–Chall score"
  ) +
  theme_minimal() +
  theme(
    text = element_text(size = 12),
    legend.position = "none"
  )



```

### Selecting less/most Readable Chunks
```{r}
library(dplyr)

# Define the chunks of interest
selected_chunks <- c("262-1", "072-1", "147-3", "011-1")

# Filter corpus_df for these chunks and select relevant columns
chunks_table <- corpus_df %>%
  filter(chunk_id %in% selected_chunks) %>%
  select(
    chunk_id, speaker, ntokens, nsentences, 
    Flesch, FleschKincaid, DaleChall, SMOG, GunningFog, text
  )

# View the table
print(chunks_table)

# Optionally, for a nicer table in RMarkdown or exporting to Word/Excel:
# library(knitr)
# kable(chunks_table, caption = "Readability Metrics for Selected Chunks")


```




### Visualizing word that are not in the Dalle-Chall word list 1995
word list obtained from: 
ReadabilityFormulas.com. (n.d.). The Dale-Chall 3,000 word list for readability formulas. ReadabilityFormulas.com. Retrieved February 26, 2026, from https://readabilityformulas.com/word-lists/the-dale-chall-word-list-for-readability-formulas/
```{r}

library(wordcloud)

# --- Prepare Dale-Chall list ---
dale_lines <- readLines("dale_chall_words.txt", encoding = "UTF-8")
dale_words <- tolower(unlist(strsplit(dale_lines, "\\s+")))
dale_words <- dale_words[dale_words != ""]

# --- Tokenize corpus without lemmatization ---
corpus_tokens <- balanced_chunks %>%
  select(speaker, text) %>%
  unnest_tokens(token, text) %>%
  # Remove punctuation, numbers, symbols, and common stopwords
  filter(
    !token %in% stop_words$word,
    !grepl("^[[:punct:][:digit:]]+$", token), 
    !is.na(token),
    token != "'s"
  ) %>%
  mutate(token = tolower(token)) %>%
  mutate(non_dale = !(token %in% dale_words))

# --- Top non-Dale-Chall words per speaker ---
non_dale_tokens <- corpus_tokens %>%
  filter(non_dale) %>%
  count(speaker, token, sort = TRUE)

# --- Separate by speaker ---
non_dale_reagan <- non_dale_tokens %>% filter(speaker == "Reagan")
non_dale_trump <- non_dale_tokens %>% filter(speaker == "Trump")
```


```{r}

library(wordcloud)

# --- Wordcloud function with custom colors ---
plot_wordcloud_colors <- function(df, colors) {
  set.seed(123)
  wordcloud(
    words = df$token,
    freq  = df$n,
    max.words = 150,
    scale = c(2, 0.3),                  # smaller words to fit more
    colors = sample(colors, length(df$token), replace = TRUE),
    random.order = FALSE
  )
}

# --- Define color palettes ---
reagan_colors <- c("#FF6666", "#FF0000", "#CC0000", "#990000", "#660000")
trump_colors  <- c("#66CCCC", "#00CCCC", "#00AAAA", "#009999", "#006666")

# --- Side-by-side plotting ---
par(mfrow = c(1,2), mar = c(0,0,0,0))  # 1 row, 2 columns, no margins

# Reagan wordcloud
plot_wordcloud_colors(non_dale_reagan, reagan_colors)

# Trump wordcloud
plot_wordcloud_colors(non_dale_trump, trump_colors)

# Reset plotting layout
par(mfrow = c(1,1))



```



### Comparison: Word Familiarity vs Word Complexity
```{r}

# DaleChall vs SMOG 

# Scatter plot: Dale-Chall vs Gunning Fog
ggplot(corpus_df, aes(x = DaleChall, y = SMOG, color = speaker)) +
  geom_point(alpha = 0.6, size = 2) +            # points for each chunk
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed") + # optional trend line
  labs(
    title = "Comparison: Word Familiarity vs Word Complexity",
    x = "Dale-Chall Score",
    y = "Smog Score",
    color = "Speaker"
  ) +
  theme_minimal(base_size = 14)

```




